### 2009-09-15 trunk/latex/paper1/contents.tex
paper

@@

Unfortunately there are few systems with known reference modularizations and,
because to obtain reference modularizations is costly, there are few empirical
studies, and most of them consider a couple of small and medium systems.

@@

\section{Software Modularization Recovery: an Overview}

Aka software architecture recovery, software architecture reconstruction,
software clustering...

It's an task of reverse engineering, as defined by Tonella \cite{Tonella2007}...

DEFINITIONS

external edge...

@@

\section{Network Theory}

@@

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to generate networks that are scale-free. In this section we present three
models that generate directed networks with built-in modular decomposition:
BCR+, CGW and LR. The first two models grow networks using a mechanism called
preferential attachment CITE Barabasi, by which nodes with many edges tend to
receive more edges in the process.

@@

\subsection{BCR plus}

The BCR model aims to model the network of hyperlinks between web pages as a
directed graph without modules \cite{Bollobas2003}. It was proven analitically
to generate scale-free networks. We have developed an extension to this model,
which we will call BCR+, that adds the concept of module. The model accepts the
following parameters:

@@

In a network with modules, one can define a MDG as a graph where the two
following conditions hold:
* each vertex represents a module in the original network;
* there is an edge from M1 to M2 in the MDG only if there is an edge from 
v1 to v2 in the original network, where v1 in M1 and v2 in M2.

The BCR+ model generates networks whose MDG is equal to G by adding vertices and
edges to an initial network until it reaches n vertices. The initial network is
isomorphic to G: it contains one vertex for each module and one edge for each
edge in G, connecting vertices whose corresponding modules are are connected.
Thus, the networks's MDG is equal to G from the very beginning.

After that, the algorithm consists of successive applications of one out of
three operations on the network: (1) adding a vertex with an outgoing edge; (2)
adding a vertex with an ingoing edge; (3) adding a edge between existing
vertices. The choice of the vertices that will be connected by a new edge,
although non-deterministic, is not fully random. The probability that a
particular vertex v is choosen, P(v), is proportional to a function of the
in-degree or of the out-degree of the node. We say that we choose a vertex
within a set V according to a function f if

  P(choose v) = f(v) / sum(v in V, f(v))

@@

* out-neighbors(v): the set of all vertices that are connected to v by an edge
starting in v.
* same-module(v): the set of all vertices that are in the same module as v (except
v itself).
* other\_modules(v): the set of all vertices that are in modules that are connected
to v's module (in the graph G) by an edge starting in v's module.

@@

\begin{figure*}
\begin{verbatim}
while the network has less than n vertices:
  choose one of the operations according to probabilities (p, q, r)
  operation 1: (add a vertex with an outgoing edge)
    choose a vertex w within V according to f(x) = din + in-degree(x)
    add a new vertex v to w's module
    add an edge from v to w
  operation 2: (add a vertex with an ingoing edge)
    choose a vertex w within V according to f(x) = dout + out-degree(x)
    add a new vertex v to w's module
    add an edge from w to v
  operation 3: (add an edge between two existing vertices)
    choose a vertex v within V according to f(x) = dout + out-degree(x)
    choose a case according to probabilities (mu, 1 - mu)
    case 1: (distinct modules)
  	choose a vertex w within (other\_modules(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
    case 2: (same module)
      choose w within (same-module(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
  end
\end{verbatim}
\end{figure*}

@@

* n, m
* Probabilities p1, p2, p3, p4, summing 1
* Natural numbers e1, e2, e3, e4
* alpha

@@

The LF model is a very flexible model that can generated weighted directed
networks with overlapping modules, that is, in which a vertex can belong to more
than one module. Unlike the previous models, this is not a growth model: all
vertices are generated at once and then the edges are added.

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at http://.

@@

Our research hypothesis is that at least one of the presented models can
synthesize networks that resemble software networks. A central issue, thus, is
how to measure similarity between networks. In order to be useful, the metric
must be able to differentiate between software and non-software networks. In
this section we present such a metric, together with an experiment that
evaluates its usefulness by applying it to both software and non-software
networks. %Then we show a classification model

\subsection{Similarity Between Networks}

@@

@@

By counting how many times each triad appears in a network, one can build a
triad concentration profile (TCP), which is a vector with 13 numbers that
summarize the local structure of the network. Figure \ref{fig:profiles} shows
the TCP for three distinct networks.

@@

% Word adjacency network: word Y follows word X => X->Y
%\begin{figure*}[!t]
%\center
%\subfigure[ref1][Software
%network]{\includegraphics[width=2.5in]{triad-jabref}}
%\hfil
%\subfigure[ref2][Non-software network]{\includegraphics[width=2.5in]{triad-lang}%
%}
%\caption{Triad profiles}
%\label{fig:profiles}
%\end{figure*}

\begin{figure}[!t]
\center
\includegraphics{tcp}

\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see Appendix A). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

\subsection{Data Set}

@@

\textbf{Non-software networks}. We have collected 66 networks from distinct
domains, such as biology, sociology, technology, and linguistics. These networks
are freely available on the Internet and have been used in previous researches.

@@

\subsection{Evaluation of the Similarity Metric}

@@

where $S$ is the set of sample software networks, and $|S|$ is the cardinality
of $S$. In this work we use the full software data set consisting of 65 software
networks.

We measured the S-score for each software network, which ranged from 0.83 to
0.98. The average S-score was 0.97 and the standard deviation, 0.03. The high
average S-score and the low standard deviation show that the metric successfully
captures common structural patterns in software networks.

@@

Then we measured the S-score for each non-software network. The majority of the
networks (X\%) have a S-score lower than 0.83, which shows that they
are ... Networks extracted from the social network Facebook have negative
S-score.

A few networks, though, showed high S-score. The network
polblogs, showing links between blogs on politics, had similarity 0.97. The
neural network of C Elegans also had a high similarity (0.88). Further
investigation is needed in order to discover why does it happen and whether
auxiliary metrics can differentiate these networks from software networks.

@@

As we have shown on the previous section, there are non-software networks with
high S-scores, hence it is impossible to build a perfect classification model,
regardless of the threshold. Nonetheless, such a model can be evaluated by its
precision and recall. Consider a data set with both software and non-software
networks. Let $S$ be the set of all software networks, and $L$ the set of all
networks that were classified as software-like. The precision of the model is

@@

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be X\%, which is too low, meaning the most software networks are
misclassified. So we chose the value 0.88, that is immediatly greater than the second greater
non-software network S-score. With this value, we have a high recall (X\%) and a
still high precision (X\%).

@@

The probabilities p, q and r were given all possible values from 0.0 up to 1.0,
in 0.2 steps, such as the sum of the probabilities was 1. Since the only events
that create vertices are those associated with probabilities p and q, we imposed
the additional restriction that p + q > 0.

For deltain and deltaout we assigned the integer numbers from 0 to 4. TODO: why

@@

Like BCR, we choose mu ranging from 0.0 to 0.6 in 0.2 steps. For the remaining
parameters we selected values from our sample of software systems. For degexp,
..., we picked the minimum, the median and the max values.

Total: 1,296 networks

@@

We have varied each of the probabilities p1, p2, p3, and p4 from 0.0 to 1.0, in
0.2 steps, with $p1 > 0$ and $p1 + p2 + p3 + p4 = 1$. Each of the parameters
$e1$, $e2$, $e3$, and $e4$ assumed the values 1, 2, 4, and 8.

% We have imposed
%the additional restriction that $2 \cdot p4\cdot e4 \ge p1 \cdot e1 + p2\cdot e2$, 
%so the number of edges created is at least twice the number of edges removed
%(otherwise 
% TODO: DOES NOT MAKE SENSE

alpha in -1, 0, 1, 10, 100, 1000. 
m in 2, 4, 8, 16, 32 (just like bcr and lfr)

@@

We used our classifier ...

Table: Model | Number of Networks | Realistic Networks | Percent \%

Show some graphs: histogram of average correlations for each model.
\ref{fig:histograms}

%\begin{figure*}[!t]
%\center
%
%\subfigure[CGW]{\includegraphics[width=1.0in]{hist-cgw}}
%
%\subfigure[BCR+]{\includegraphics[width=1.0in]{hist-bcr}}
%
%\subfigure[LF]{\includegraphics[width=1.0in]{hist-lf}}
%
%\caption{Triad profiles}
%\label{fig:histograms}
%\end{figure*}

All models produce networks that resemble software networks.  For some
parameters, though, the networks are not realistic.

\subsection{Parameter-Based Classification Models}

1R

Naive Bayes

\subsection{Homogeinity}

Pick realistic networks from a model. Are they similar to each other? (see
standard deviation) In other words: do the parameters make a difference?

Are they similar to networks generated by the other models? In other words: are
the models equivalent?

\section{Threats to validity}

% External validity (EV): can it be generalized?

%Structural information isn't enough for a expert do produce a
%decomposition (s/he may use data such as names and external documentation). 
%
%Even when considering only structural information, is it true
%that experts would find a decomposition similar to the reference decomposition
%imposed by the model?

We generated only one network for each set of parameters. (as we've shown, some
parameters are redundant as they do not change significantly the realism)

Some clustering algorithms use weights and they weren't studied here.

EV: We've only studied 65 systems, which is not that much.

We only studied object-oriented systems implemented in Java. Maybe the results
would be different if we studied systems implement in other languagens or using
other paradigms. The choice of a particular technique for extracting
dependencies (static analysis) may also have impact on the structure of the
networks.

% Koschke says that experts decompositions vary by no more than 80? percent.

>>>>

TODO: sw clustering is a task of reverse engineering \cite{Tonella2007}.

@@

Unfortunately there are few systems with known reference modularizations and,
because it is costly to obtain reference modularizations, there are few
empirical studies, and most of them consider a couple of small and medium
systems.

@@

%\section{Software Modularization Recovery: an Overview}
%Aka software architecture recovery, software architecture reconstruction,
%software clustering...
%It's an task of reverse engineering, as defined by Tonella \cite{Tonella2007}...
%DEFINITIONS
%external edge...

@@

\section{Software Networks}

Define: Component. Component dependency network. Directed unweighted graph.

Clustering based on dependencies.

@@

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to generate networks that are scale-free. In this section we present three
models that generate directed networks with built-in modular decomposition:
BCR+, CGW and LR. 
%The first two models grow networks using a mechanism called
%preferential attachment CITE Barabasi, by which nodes with many edges tend to
%receive more edges in the process.

@@

\subsection{BCR+}

The BCR model aims to model the network of hyperlinks between web pages as a
directed graph without modules \cite{Bollobas2003}. It was proven analitically
to generate scale-free networks. We have developed an extension to this model,
called BCR+, that adds modules to the construction of the network. The model
accepts the following parameters:

@@

In a network with modules, one can define its module dependency graph (MDG) as a
graph where the two following conditions hold:

\begin{itemize}
\item each vertex represents a module in the original network; 
\item there is an edge from M1 to M2 in the MDG only if there is an edge from v1
to v2 in the original network, where v1 in M1 and v2 in M2.
\end{itemize}

The BCR+ model generates networks whose MDG is equal to G by adding vertices and
edges to an initial network until it reaches n vertices. The initial network is
isomorphic to G: it contains one vertex for each module and one edge for each
edge in G, connecting vertices whose corresponding modules are are connected.
Thus, the MDG for the generated network is equal to G from the very beginning.

After that, the algorithm consists of successive applications of one out of
three operations on the network: (1) adding a vertex with an outgoing edge; (2)
adding a vertex with an ingoing edge; (3) adding a edge between existing
vertices. In operation 3, the edge can connect either vertices that belong
to the same module or vertices in distinct modules.

The choice of the vertices that will be connected by a new edge, although
non-deterministic, is not fully random. The probability that a particular vertex
v is choosen, P(v), is proportional to a function of the in-degree or the
out-degree of the node. We say that we choose a vertex within a set of vertices
S according to a function f if

$$
  P(choose v) ~=~ \frac{ f(v) }{ \sum_{v \in S, f(v)} }.
$$

@@

\begin{itemize}

\item out-neighbors(v): the set of all vertices that are connected to the vertex v by an edge
starting in v.

\item same-module(v): the set of all vertices that are in the same module as the
vertex v, except for v itself and vertices in out-neighbors(v).

\item other\_modules(v): the set of all vertices that are in modules that are
connected to v's module (in the graph G) by an edge starting in v's module,
except for vertices in out-neighbors(v).

\end{itemize}

@@

\begin{figure*}
\begin{verbatim}
while the number of nodes in the network is less than n:
  choose one of the operations according to probabilities (p, q, r)
  operation 1: (add a vertex with an outgoing edge)
    choose a vertex w within V according to f(x) = din + in-degree(x)
    add a new vertex, v, to w's module
    add an edge from v to w
  operation 2: (add a vertex with an ingoing edge)
    choose a vertex w within V according to f(x) = dout + out-degree(x)
    add a new vertex, v, to w's module
    add an edge from w to v
  operation 3: (add an edge between two existing vertices)
    choose a vertex v within V according to f(x) = dout + out-degree(x)
    choose a case according to probabilities (mu, 1 - mu)
    case 1: (distinct modules)
  	  choose a vertex w within other\_modules(v) according to f(x) = din + in-degree(x)
      add an edge from v to w
    case 2: (same module)
      choose w within same-module(v) according to f(x) = din + in-degree(x)
      add an edge from v to w
\end{verbatim}
\end{figure*}

@@

\begin{itemize}
\item Number of vertices, n;
\item Number of modules, m;
\item Probabilities p1, p2, p3, p4, summing 1;
\item Natural numbers e1, e2, e3, e4;
\item Number alpha, with alpha >= -1
\end{itemize}

@@

The LF model \cite{Lancichinetti2009} is a very flexible model that can
generated weighted directed networks with overlapping modules, that is, in which
a vertex can belong to more than one module. Unlike the previous models, this is
not a growth model: all vertices are generated at once and then the edges are
added.

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at http://. It accepts the following
parameters:

\begin{itemize}
\item number of nodes, n;
\item average in-degree, k;
\item maximum in-degree, maxk;
\item mixing parameter, mu;
\item exponent for the degree distribution, t1;
\item exponent for the module size distribution, t2;
\item minimum module size, minm;
\item maximum module size, maxm;
\end{itemize}

@@

Our research hypothesis is that at least one of the presented models can
synthesize networks that resemble software networks. A central issue, thus, is
how to measure similarity between networks. In order to be useful, a similarity
metric must be able to differentiate between software and non-software networks.
In this section we present such a metric, together with an experiment that
evaluates its usefulness by applying it to both software and non-software
networks. %Then we show a classification model

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Similarity Between Networks}

@@

CITE igraph

@@

By counting how many times each triad appears in a network, one can build a
triad concentration profile (TCP), which is a vector with 13 numbers that
summarize the local structure of the network. Figure \ref{fig:profiles} shows
the TCP for networks from distinct domains.

@@

% Word adjacency network: word Y follows word X => X->Y
\begin{figure}[!t]
\center
\includegraphics{tcp}
\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see Appendix A). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Data Set}

@@

\textbf{Non-software networks}. We have collected 66 networks from distinct
domains, such as biology, sociology, technology, and linguistics. These networks
are freely available on the Internet and have been used in previous researches.
CITE

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation of the Similarity Metric}

@@

where $S$ is the set of sample software networks, and $|S|$ is the number of
networks in $S$. In this work we use the full software data set consisting of 65
software networks as our sample.

We measured the S-score for each software network, which ranged from 0.83 to
0.98. The average S-score was 0.97 and the standard deviation, 0.03. The high
average S-score and the low standard deviation show that the metric successfully
characterizes software networks by capturing common structural patterns

@@

Then we measured the S-score for each non-software network. The majority of the
networks (X\%) had a S-score lower than 0.83, which is the lowest S-score for
software networks in the sample. Some social networks even showed negative
S-score, meaning that they are very different from software networks.

Two networks, though, showed high S-score, above 0.83. The network polblogs,
showing links between blogs on politics, showed S-score 0.97. The neural network
of the worm C.  Elegans also showed a high S-score (0.88). Further investigation
is needed in order to discover the reasons behind these high S-scores and
whether auxiliary metrics can differentiate these networks from software
networks.

@@

As we have shown on the previous section, there are non-software networks with
high S-scores, hence it is impossible to build a perfect classification model,
regardless of the threshold. Nonetheless, such a model can be evaluated by its
precision and recall. Consider a data set with both software and non-software
networks. Let $S$ be the set of all software networks, and $L$ the set of all
networks that were classified by the model as software-like. The precision of
the model is

@@

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be X\%, which is too low, meaning the most software networks are
misclassified. So we chose the value 0.88, that is immediatly greater than the
second greater non-software network S-score. With this value, we have both a
high recall (X\%) and a high precision (X\%).

@@

The probabilities p, q and r were given all possible values from 0.0 up to 1.0,
in 0.2 steps, such as the sum of the probabilities was 1. Since the only events
that create vertices are those associated with probabilities p and q, we imposed
the additional restriction that $p + q > 0$.

For deltain and deltaout we assigned integer numbers ranging from 0 to 4. TODO: why

@@

Like BCR+, we choose mu ranging from 0.0 to 0.6 in 0.2 steps. For the remaining
parameters, the following values were used:

\begin{itemize}
  \item t1 in 2.18, 2.70, 3.35 (degree exponent distribution)
  \item t2 in 0.76, 0.99, 1.58
  \item k in 5, 10, 15, 25 
  \item maxk in 58, 157, 482
  \item minc 1, 10, 273
\end{itemize}

In order to choose these values, we analyzed software networks from our sample
with approximately 500 to 2,000 nodes, so no network was much bigger or much
smaller than the synthetic networks. We computed the exponents for the degree
and module size distributions using KOLMOGOROV-SMIRNOV, GOODNESS-OF-FIT., and
then chose the minimum, median and maximum values. For k, maxk, minc the values
extracted from the network were divided by the number of nodes in the network
and then multiplied by 1000. For maxk and minc we chose the minimum, median and
maximum values. For k .... The parameter maxc was left unbound to avoid
impossible combinations of minc, maxc, and t2.

In total, 1,296 networks where synthesized.

@@

The parameters values were chosen as such:

\begin{itemize}
\item $p1, p2, p3, p4 \in {0.0, 0.2, 0.4, 0.6, 0.8, 1.0}$, with $p1 > 0$ and $p1
  + p2 + p3 + p4 = 1$;
\item $e1, e2, e3, e4 \in {1, 2, 4, 8}$;
\item $\alpha \in {-1, 0, 1, 10, 100, 1000}$;
\item number of modules: $m \in {2, 4, 8, 16, 32}$.
\end{itemize}

Notice that ei has no effect when pi is zero; in this case ei was just set to
zero.

@@

Each synthesized network was classified as software-like or non software-like,
using the classification model presented in section XXX. The results are
summarized in Table X.

\begin{table}
Table: Model | Realistic Networks | Percent \%
\end{table}

All models synthesized both software-like and non software-like networks. The
proportion of software-like networks was greater than X\% for all models,
discarding the possibility that this result was obtained by pure chance.

%Show some graphs: histogram of average correlations for each model.
%\ref{fig:histograms}
%\begin{figure*}[!t]
%\center
%
%\subfigure[CGW]{\includegraphics[width=1.0in]{hist-cgw}}
%
%\subfigure[BCR+]{\includegraphics[width=1.0in]{hist-bcr}}
%
%\subfigure[LF]{\includegraphics[width=1.0in]{hist-lf}}
%
%\caption{Triad profiles}
%\label{fig:histograms}
%\end{figure*}

Of course, this result is of little practical value unless there is a
relationship between parameter values and S-score. For the purpose of this
research, it is important to know which values are more likely to lead to
software-like networks.

The algorithm 1R from data mining was used to help discover such relationship.
It analyzes the parameters and the classification of each network and finds a
rule that relates the value of one single parameter with the classification. We
are interested in rules for networks classified as software-like and in the
precision of these rules. The rules found by 1R are shown in Table X.

Model | Rule | Precision

Despite its simplicity, 1R found rules with high precision. These rules allow
one to know if a particular combination of parameter values is likely to lead to
software-like networks.

%\subsection{Parameter-Based Classification Models}
%
%1R

%Naive Bayes

%\subsection{Homogeinity}
%
%Pick realistic networks from a model. Are they similar to each other? (see
%standard deviation) In other words: do the parameters make a difference?
%
%Are they similar to networks generated by the other models? In other words: are
%the models equivalent?

%\section{Threats to validity}
%
%% External validity (EV): can it be generalized?
%
%%Structural information isn't enough for a expert do produce a
%%decomposition (s/he may use data such as names and external documentation). 
%%
%%Even when considering only structural information, is it true
%%that experts would find a decomposition similar to the reference decomposition
%%imposed by the model?
%
%We generated only one network for each set of parameters. (as we've shown, some
%parameters are redundant as they do not change significantly the realism)
%
%Some clustering algorithms use weights and they weren't studied here.
%
%EV: We've only studied 65 systems, which is not that much.
%
%We only studied object-oriented systems implemented in Java. Maybe the results
%would be different if we studied systems implement in other languagens or using
%other paradigms. The choice of a particular technique for extracting
%dependencies (static analysis) may also have impact on the structure of the
%networks.
%
%% Koschke says that experts decompositions vary by no more than 80? percent.
